Title

### 2013-03-05 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
improve formal external_forces section
merging comments on Liouville's theorem

\section{Implications of Liouville's Theorem}

One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
%TODO find a good reference for this

To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
Using \ref{eq:ag_original} it is easy to show that
\begin{gather}
  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
  \label{eq:define_big_gamma}
  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
\end{gather}
This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.

Not only is this an interesting result, but it can be both instructive and computationally useful.
By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, evaulating at some time (presumably $t=0$) and substituting into \ref{eq:ag_original}.
\begin{subequations} \label{eq:ag_big_gamma}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
\end{gather}
\end{subequations}
these four equations now contain all of the information previously contained in six.
Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
This will become more apparent when in future sections
%TODO include internal reference
it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

>>>>

\section{Implications of Liouville's Theorem} \label{sec:liouville}

One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
%TODO find a good reference for this
In the AG model the component phase-space products are simply $\sigma_{\alpha} \eta_{\alpha}$ where as before $ \alpha = \smallT, z$.

In fact it is easy to show that the AG model explicitly obeys the Liouville Theorem.
To do so, consider the time derivative of these component phase-space products.
Using \ref{eq:ag_original} it is easy to show that
\begin{equation}
  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0
\end{equation}
and therefore we can define the per-simulation constants $\Gamma_{\alpha}$ as
\begin{equation} \label{eq:define_big_gamma}
  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} = \text{Constant .}
\end{equation}

Not only is this an interesting result, but it can be both instructive and computationally useful.
By evaluating $\Gamma_{\alpha}$ at some time (presumably $t=0$) and rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, we can rewrite \ref{eq:ag_original} as 
\begin{subequations} \label{eq:ag_big_gamma}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) \, \text{.} \label{eq:ag_big_gamma_gamma}
\end{gather}
\end{subequations}
These four equations now contain all of the information previously contained in six (\ref{eq:ag_original}).
Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
This will become more apparent when in future sections
%TODO include internal reference
it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.
